---
title: "Getting Started with openneuro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with openneuro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why openneuro?

[OpenNeuro](https://openneuro.org) is the largest open neuroimaging data
repository, hosting thousands of BIDS-formatted datasets spanning MRI, EEG,
MEG, and more. Accessing these datasets programmatically — searching by
modality, downloading specific subjects, managing local caches — typically
requires juggling GraphQL queries, S3 URLs, and file-system bookkeeping.

The `openneuro` package handles all of that so you can focus on your analysis.
With a few function calls you can search the catalogue, inspect dataset
contents, and download exactly the files you need.

## Load the package

```{r load-package}
library(openneuro)
```

## Find datasets

Search the OpenNeuro catalogue. Results come back as a tibble:

```{r search-datasets, eval=FALSE}
ds <- on_search(limit = 10)
ds[, c("id", "name", "n_subjects")]
```

```
#> # A tibble: 10 x 3
#>    id        name                              n_subjects
#>    <chr>     <chr>                                  <int>
#>  1 ds000001  Balloon Analog Risk-taking Task           16
#>  2 ds000002  Classification Learning                   17
#>  3 ds000003  Rhyme Judgment                            13
#>    ...
```

You can filter by modality:

```{r search-modality, eval=FALSE}
mri <- on_search(modality = "MRI", limit = 25)
head(mri$id)
```

```
#> [1] "ds000001" "ds000002" "ds000003" "ds000005" "ds000006" "ds000007"
```

Inspect a single dataset:

```{r inspect-dataset, eval=FALSE}
meta <- on_dataset("ds000001")
meta
```

## Inspect versions and files

Each dataset has one or more snapshots (versioned releases):

```{r list-snapshots, eval=FALSE}
snaps <- on_snapshots("ds000001")
snaps
```

```
#> # A tibble: 3 x 3
#>   tag    created                    size
#>   <chr>  <chr>                      <chr>
#> 1 1.0.0  2018-02-02T00:00:00.000Z   2.1 GB
#>   ...
```

List files from the latest snapshot:

```{r list-files, eval=FALSE}
files <- on_files("ds000001")
head(files)
```

```
#> # A tibble: 6 x 3
#>   filename                   size     directory
#>   <chr>                      <chr>    <lgl>
#> 1 dataset_description.json   1.2 kB   FALSE
#> 2 participants.tsv           800 B    FALSE
#>   ...
```

## Download data

Download specific files to your local cache:

```{r download-files, eval=FALSE}
res <- on_download(
  id = "ds000001",
  files = c("dataset_description.json", "participants.tsv"),
  quiet = FALSE
)
res
```

Download by subject IDs — the package normalises bare numbers and `"sub-"`
prefixed IDs automatically:

```{r download-subjects, eval=FALSE}
res_sub <- on_download(
  id = "ds000001",
  subjects = c("01", "02"),
  include_derivatives = FALSE
)
```

Use a regex to select subjects by pattern:

```{r download-regex, eval=FALSE}
res_rx <- on_download(
  id = "ds000001",
  subjects = regex("sub-0[1-5]")
)
```

### Manage the cache

Inspect and clean up your local cache:

```{r cache-utils, eval=FALSE}
on_cache_info()
on_cache_list()

# Clear one dataset from cache
on_cache_clear("ds000001", confirm = FALSE)
```

## Discover derivatives

Many OpenNeuro datasets have community-contributed derivative outputs
(e.g. fMRIPrep, MRIQC). You can list them:

```{r list-derivatives, eval=FALSE}
derivs <- on_derivatives("ds000001")
derivs[, c("dataset_id", "pipeline", "source")]
```

Inspect available output spaces for a derivative:

```{r inspect-spaces, eval=FALSE}
spaces <- on_spaces(derivs[1, ])
spaces
```

Download derivative outputs for specific subjects and spaces:

```{r download-derivatives, eval=FALSE}
on_download_derivatives(
  dataset_id = "ds000001",
  pipeline = "fmriprep",
  subjects = c("01", "02"),
  space = "MNI152NLin2009cAsym"
)
```

## Next steps

- `vignette("openneuro-fmriprepper-e2e")` — a full end-to-end workflow that
  downloads an OpenNeuro dataset and processes it with fMRIPrep via
  `fmriprepper`.
- `?on_search`, `?on_download`, `?on_derivatives` — detailed function
  documentation.
