---
title: "End-to-End: OpenNeuro Download + fmriprepper Run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-End: OpenNeuro Download + fmriprepper Run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette walks through a complete analysis pipeline: downloading a
BIDS dataset from OpenNeuro and processing it with
[fMRIPrep](https://fmriprep.org) using the
[fmriprepper](https://github.com/bbuchsbaum/fmriprepper) package. By the end
you'll have preprocessed fMRI data ready for statistical analysis.

## Prerequisites

You'll need:

- The `openneuro` package (this package).
- The `fmriprepper` package:
  `remotes::install_github("bbuchsbaum/fmriprepper")`
- A working fMRIPrep runtime — either Docker or Apptainer/Singularity with an
  fMRIPrep container image.

## Reading the fluent `$` syntax

`fmriprepper` uses an R6 "fluent" API. In chains like:

```r
plan <- fmriprep()$engine_singularity(...)$bids(...)$out(...)
```

each `$method(...)` call configures the same plan object and returns it, so the
next `$...` keeps building that same object.

If you prefer a non-chained style, this is equivalent:

```r
plan <- fmriprep()
plan$engine_singularity(image = "/path/to/fmriprep.sif", runtime = "apptainer")
plan$bids(dataset_dir)
plan$out(deriv_dir)
plan$work(path(work_dir, "sub-{subject}"))
plan$participant_labels(c("01"))
plan$nprocs(16)
plan$omp_nthreads(8)
plan$mem_mb(32000)
```

## Step 1: Download an OpenNeuro dataset

First, set up your project directory structure and download the dataset. The
`on_download()` function supports multiple backends; here we use DataLad for a
repository-style clone (fall back to `backend = "https"` if DataLad is not
installed).

```{r download-dataset, eval=FALSE}
library(openneuro)
library(fs)

dataset_id <- "ds000001"
project_root <- path_expand("~/data/openneuro-fmriprepper")
dataset_dir  <- path(project_root, dataset_id)
deriv_dir    <- path(project_root, "derivatives", "fmriprep")
work_dir     <- path(project_root, "work")

dir_create(project_root, recurse = TRUE)

download_result <- on_download(
  id       = dataset_id,
  dest_dir = dataset_dir,
  use_cache = FALSE,
  backend  = "datalad",
  quiet    = FALSE
)

print(download_result)
stopifnot(file_exists(path(dataset_dir, "dataset_description.json")))
```

Adjust `project_root` to wherever you'd like to store data on your machine.

## Step 2: Build an fMRIPrep plan

The `fmriprepper` package provides a builder-style interface for constructing
fMRIPrep commands. You chain methods to configure the container engine, input
paths, resource limits, and participant selection.

**Using Apptainer/Singularity:**

```{r plan-singularity, eval=FALSE}
library(fmriprepper)

plan <- fmriprep()$
  engine_singularity(
    image   = "/path/to/fmriprep.sif",
    runtime = "apptainer"
  )$
  bids(dataset_dir)$
  out(deriv_dir)$
  work(path(work_dir, "sub-{subject}"))$
  participant_labels(c("01"))$
  nprocs(16)$
  omp_nthreads(8)$
  mem_mb(32000)
```

**Using Docker** (replace the `engine_singularity()` call):

```{r plan-docker, eval=FALSE}
plan <- fmriprep()$
  engine_docker(image = "nipreps/fmriprep:latest")$
  bids(dataset_dir)$
  out(deriv_dir)$
  work(path(work_dir, "sub-{subject}"))
```

Replace `/path/to/fmriprep.sif` with the actual path to your Singularity
image.

## Step 3: Validate and run

Before executing, validate that the runtime environment and BIDS directory are
set up correctly:

```{r validate, eval=FALSE}
plan$validate_runtime(strict = TRUE, verbose = TRUE)
plan$validate_bids_light(strict = TRUE, verbose = TRUE)
```

Preview the commands that will be generated:

```{r preview-commands, eval=FALSE}
cmds <- plan$commands(batch = TRUE, skip_completed = TRUE)
cat(cmds, sep = "\n")
```

When you're satisfied, execute:

```{r execute, eval=FALSE}
plan$run(batch = TRUE, skip_completed = TRUE, parallel = 1)
```

## Tips

- Change `dataset_id` and `participant_labels()` to match your study.
- For HPC workflows, use `plan$write_manifest()` and
  `plan$write_slurm_array()` from `fmriprepper`.
- See `vignette("openneuro")` for more on dataset search and download options.
